Classical Mechanics Assignment Two

### 2015-10-09 classical/latex/a2.tex
Push errthang.

\title{Classical Mechanics Assignment One}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

@@

\begin{equation}
  m\frac{dv}{dt} = -mg + cv^2\\
\end{equation}
This situation corresponds to either upward or downward motion. If $\sqrt(\frac{mg}{c}) < v$, then the object will constantly feel upward motion. If $\sqrt(\frac{mg}{c}) > v$, then the object will at first feel deceleration until v is so negative that $\sqrt(\frac{mg}{c}) < v$, at which point it will begin to accelerate until this is no longer true. It will oscillate around the negative v where $\sqrt(\frac{mg}{c}) > v$ indefinitely.

@@

@@

@@

\begin{align}
  m\frac{dv}{dt} = F_f - cv^2\\
  m\frac{dv}{F_f - cv^2} = dt\\
  \int_{v0}^v m\frac{dv}{F_f - cv^2} = \int dt\\
  \int_{v0}^v \frac{mdv}{F_f - cv^2} = t\\
  \frac{m*\tanh^{-1}(\frac{\sqrt{c}v}{\sqrt{F}})}{\sqrt{cF}} - \frac{m*\tanh^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})}{\sqrt{cF}} = t\\
  \tanh^{-1}(\frac{\sqrt{c}v}{\sqrt{F}}) - \tanh^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}}) = \frac{t\sqrt{cF}}{m}\\
  \tanh^{-1}(\frac{\sqrt{c}v}{\sqrt{F}}) = \frac{t\sqrt{cF}}{m} + \tanh^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})\\  
  \tanh(\frac{t\sqrt{cF}}{m} + \tanh^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})) = \frac{\sqrt{c}v}{\sqrt{F}}\\
  \sqrt{\frac{F}{c}}\tanh(\frac{t\sqrt{cF}}{m} + \tanh^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})) = v(t)\\    
\end{align}

@@

This was a harder problem, but also intersting. We had to use the basic technique of physics, finding a differential equation to model a system and solving for velocity (or usually position), then use a computer to plot this solution, since it is difficult to interpret otherwise. I learned that from very simple initial forces (a constant force and a velocity squared force, both opposing motion), the equation for velocity can become complex (this one involved a hyperbolic tangent).

@@

When falling back down:\\
\begin{align}
  m\frac{dv}{dt} = -mg + cv^2\\
  m\frac{dv}{dy}\frac{dy}{dt} = -mg + cv^2\\
  m\frac{dv}{dy}v = -mg + cv^2\\
  m\frac{dv}{-mg + cv^2}v = dy\\
  \int_{v_0}^0 m\frac{dv}{-mg + cv^2}v = \int dy\\
  \int_{v_0}^0 m\frac{mv}{-mg + cv^2}dv = -y_{max}\\
  \int_{v_0^2}^0 \frac{mv}{-mg + cu}\frac{du}{2v} = -y_{max}\\
  \frac{1}{2} \int_{v_0^2}^0 \frac{m}{-mg + cu} du = -y_{max}\\
  \frac{m \log(-gm)}{2c} - \frac{m \log(cv_0^2 - gm)}{2c} = -y_{max}\\
  \frac{m}{2c} \log(\frac{gm}{gm - cv_0^2}) = -y_{max}\\
\end{align}

\begin{align}
  \frac{m}{2c} \log(\frac{gm}{cv_0^2 + gm}) = -\frac{m}{2c} \log(\frac{gm}{gm - cv_0^2})\\
  \log(\frac{gm}{cv_0^2 + gm}) = -\log(\frac{gm}{gm - cv_0^2})\\
  \log(\frac{gm}{cv_0^2 + gm}) = \log(\frac{gm - cv_0^2}{gm})\\
  \frac{gm}{cv_0^2 + gm} = \frac{gm - cv_0^2}{gm}\\
  \frac{g^2m^2}{cv_0^2 + gm} = gm - cv_0^2\\    
\end{align}

@@

>>>>

\title{Classical Mechanics Assignment Two}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

@@

\begin{equation}
  m\frac{dv}{dt} = -mg + cv^2\\
\end{equation}
This situation corresponds to either upward or downward motion. If $\sqrt{\frac{mg}{c}} < v$, then the object will constantly feel upward motion. If $\sqrt{\frac{mg}{c}} > v$, then the object will at first feel deceleration until v is so negative that $\sqrt{\frac{mg}{c}} < v$, at which point it will begin to accelerate until this is no longer true. It will oscillate around the negative v where $\sqrt{\frac{mg}{c}} > v$ indefinitely.

@@

This solution makes sense. It has positive sign, which makes sense, since positive sign corresponds to downward motion. Alpha is inversely proportional to terminal velocity, and mass is proportional.

@@

This solution makes sense. Terminal velocity is proportional to mass and inversely proportional to alpha, as expected. For this particular system, where force is highly velocity-dependent, the terminal velocity grows a lot slower than in the above case, where force is less strongly velocity-dependent.

@@

\begin{align}
  m\frac{dv}{dt} = -F_f - cv^2\\
  m\frac{dv}{-F_f - cv^2} = dt\\
  \int_{v0}^v m\frac{dv}{-F_f - cv^2} = \int dt\\
  \int_{v0}^v \frac{mdv}{F_f + cv^2} = -t\\
  \frac{m*\tan^{-1}(\frac{\sqrt{c}v}{\sqrt{F}})}{\sqrt{cF}} - \frac{m*\tan^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})}{\sqrt{cF}} = -t\\
  \tan^{-1}(\frac{\sqrt{c}v}{\sqrt{F}}) - \tan^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}}) = \frac{-t\sqrt{cF}}{m}\\
  \tan^{-1}(\frac{\sqrt{c}v}{\sqrt{F}}) = \frac{-t\sqrt{cF}}{m} + \tan^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})\\  
  \tan(\frac{-t\sqrt{cF}}{m} + \tan^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})) = \frac{\sqrt{c}v}{\sqrt{F}}\\
  \sqrt{\frac{F}{c}}\tan(\frac{-t\sqrt{cF}}{m} + \tan^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})) = v(t)\\    
\end{align}

@@

This was a harder problem, but also intersting. We had to use the basic technique of physics, finding a differential equation to model a system and solving for velocity (or usually position), then use a computer to plot this solution, since it is difficult to interpret otherwise. I learned that from very simple initial forces (a constant force and a velocity squared force, both opposing motion), the equation for velocity can become complex (this one involved a tangent function).

@@

When falling back down:\\
\begin{align}
  m\frac{dv}{dt} = -mg + cv^2\\
  m\frac{dv}{dy}\frac{dy}{dt} = -mg + cv^2\\
  m\frac{dv}{dy}v = -mg + cv^2\\
  m\frac{dv}{-mg + cv^2}v = dy\\
  \int_{0}^{vf} m\frac{dv}{-mg + cv^2}v = \int dy\\
  \int_{0}^{vf} m\frac{mv}{-mg + cv^2}dv = -y_{max}\\
  \int_{0}^{vf} \frac{mv}{-mg + cu}\frac{du}{2v} = -y_{max}\\
  \frac{1}{2} \int_{0}^{vf} \frac{m}{-mg + cu} du = -y_{max}\\
  \frac{m \log(cv_f^2 - gm)}{2c} - \frac{m \log(-gm)}{2c} = -y_{max}\\
  \frac{m}{2c} (\log(cv_f^2 - gm) - \log(-gm)) = -y_{max}\\
  \frac{m}{2c} \log(\frac{cv_f^2 - gm}{-gm}) = -y_{max}\\
  \frac{m}{2c} \log(\frac{gm - cv_f^2}{gm}) = y_{max}\\
\end{align}

@@

\begin{align}
  \frac{m}{2c} \log(\frac{gm}{cv_0^2 + gm}) = \frac{m}{2c} \log(\frac{gm - cv_f^2}{gm})\\
  \frac{gm}{cv_0^2 + gm} = \frac{gm - cv_f^2}{gm}\\
  \frac{(gm)^2}{cv_0^2 + gm} = gm - cv_f^2\\
  cv_f^2 = gm - \frac{(gm)^2}{cv_0^2 + gm}\\
  v_f^2 = \frac{gm}{c} - \frac{(gm)^2}{c^2v_0^2 + cgm}\\
  v_f^2 = v_{term}^2 - \frac{v_{term}^4}{v_0^2 + v_{term}^2}\\
  v_f^2 = \frac{v_{term}^2(v_0^2 + v_{term}^2) - v_{term}^4}{v_0^2 + v_{term}^2}\\
  v_f^2 = \frac{v_{term}^2v_0^2}{v_0^2 + v_{term}^2}\\      
\end{align}
